---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```
[![CRAN status](https://www.r-pkg.org/badges/version/sdcSpatial)](https://cran.r-project.org/package=sdcSpatial)
[![R build status](https://github.com/edwindj/sdcSpatial/workflows/R-CMD-check/badge.svg)](https://github.com/edwindj/sdcSpatial/actions)
[![](https://cranlogs.r-pkg.org/badges/sdcSpatial)](https://cran.r-project.org/package=sdcSpatial)
[![status](https://tinyverse.netlify.com/badge/sdcSpatial)](https://CRAN.R-project.org/package=sdcSpatial)

[![Mentioned in Awesome Official Statistics](https://awesome.re/mentioned-badge.svg)](http://www.awesomeofficialstatistics.org)

# sdcSpatial

_Publishing a raster density map can reveal sensitive values_.
`sdcSpatial` is an opensource R package for creating spatial density (raster) 
maps from point data while protecting the privacy of individual observations. 

`sdcSpatial` offers a `sdc_raster` class that allows to:

- find out which locations are considered sensitive / unsafe for publishing: `plot_sensitive`, `is_sensitive`, `sensitivity_score`.
- apply protection methods that reduce sensitiviy and enhance spatial patterns: `protect_smooth`, `protect_quadtree`.
- remove sensitive locations: `remove_sensitive`.
- extract relative as well as absolute density `raster`s that can be used with visualisation packages, such as `tmap` and `leaflet`: `x$value$mean`, `x$value$sum`, `mean`, `sum`.

## Installation

```r
install.packages("sdcSpatial")
```

To install the current development version of  `sdcSpatial`  with `devtools`

```r
remotes::install_github("edwindj/sdcSpatial")
```

## Example

```{r example, code=readLines("./example/protect_smooth.R")}
```
