---
title: "Introduction sdcSpatial"
subtitle: "Privacy protected density maps"
output: 
  rmarkdown::html_vignette:
    df_print: kable
    fig_width: 6
vignette: >
  %\VignetteIndexEntry{Introduction sdcSpatial: privacy protected density maps}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  out.width = "100%"
)

# bref <- c(citation(), citation("raster"))
# writeLines(toBibtex(bref), "software.bib")

# preloading
library(sdcSpatial)
```
Plotting data on a map is a popular and helpful tool to analyse spatial data. 
In R it is easy to make spatial plots using e.g. spatial packages, `ggplot2`, 
`tmap`, `mapview` or `leaflet`. 
However when plotting the spatial distribution of a 
sensitive variable, you may unwantedly reveal a sensitive value of an 
individual observation, e.g. the income of a person. Adjusting data or output
to 

`sdcSpatial` contains  functions to create spatial distribution maps, 
assess the risk of disclosure on a location and to suppress or adjust 
revealing values at certain locations.

```{r}
library(sdcSpatial)
library(raster)
```
## `raster`

`sdcSpatial` builds heavily upon the `raster` package.

## Assessment

```{r}
data("enterprises")
class(enterprises)
summary(enterprises@data)
```
```{r}
production <- sdc_raster(enterprises, enterprises$production, r = 500)
print(production)
```

## Protection methods

```{r}
plot(production)
```
```{r}
plot(remove_sensitive(production))
```


```{r}

```

