---
title: "Introduction sdcSpatial"
subtitle: "Privacy protected density maps"
output: 
  rmarkdown::html_vignette:
    df_print: kable
    fig_width: 6
vignette: >
  %\VignetteIndexEntry{Introduction sdcSpatial: privacy protected density maps}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  out.width = "100%"
)

# bref <- c(citation(), citation("raster"))
# writeLines(toBibtex(bref), "software.bib")

# preloading
library(sdcSpatial)
```
Plotting data on a map is a popular and helpful tool to analyse spatial data. Increasingly spatial point
data is becoming available resulting in more and more spatial plots that not only are used for analytical purposes, but also for publication and communication. 
In R it is easy to make spatial plots using e.g. the rspatial packages, `ggplot2`, `tmap`, `mapview` or `leaflet`. However when the plot is meant to show the spatial distribution of a sensitive variable, it is possible that unwantedly a sensitive value of an individual person or enterprise is revealed. 
It is desirable to make a plot that contains spatial patterns (if any) 
but protects the privacy of the observed population as much as possible.

`sdcSpatial` contains  functions to create spatial distribution maps, assess the risk of disclosure on a location and to suppress or adjust revealing values at certain locations.

```{r}
library(sdcSpatial)
library(raster)
```
## `raster`

`sdcSpatial` builds heavily upon the `raster` package.

## Assessment

```{r}
data("enterprises")
class(enterprises)
summary(enterprises@data)
```
```{r}
production <- sdc_raster(enterprises, enterprises$production, r = 500)
print(production)
```

## Protection methods

```{r}
plot(production)
```
```{r}
plot(remove_sensitive(production))
```


```{r}

```

