---
title: "Privacy protecting maps"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Privacy protecting maps}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
# preloading
library(sdcSpatial)
```
Plotting data on a map is a popular and helpful tool to analyse spatial data. Increasingly spatial point
data is becoming available [ref], resulting in more and more spatial plot  
that not only are used for analytical purposes, but also for publication and communication. 
In R [ref] it is easy to make spatial plots using e.g. the rspatial packages, ggplot, tmap, mapview or leaflet. However when the plot is meant to show the spatial distribution of a sensitive variable, it is possible that unwantedly a sensitive value of an individual person or enterprise is revealed. 
It is desirable to make a plot that contains spatial patterns (if any) but protects the privacy of the 
observed population as much as possible.

`sdcSpatial` contains  functions to create spatial distribution maps, assess the risk of disclosure on a location and to suppress or adjust revealing values at certain locations.

```{r}
library(sdcSpatial)
```

## Methods

### `calc_block`

### `smooth_raster`

### `quadtree`
